---
title: "week_2_607"
author: "Justin Herman"
date: "February 8, 2018"
output: html_document

---

### Load in packages 

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
library(RMySQL)
library(ggplot2)
library(dplyr)

```
### SQL code to create a db/table in SQL
+ The code should not be executed, it needs to be run in sql
``````{r}
#DROP TABLE IF EXISTS Ratings;

#Create database Movies_db;
#use Movies_db;
#CREATE TABLE Ratings (
#`Person` VARCHAR(100) PRIMARY KEY,
#`Gladiator`  int  NOT NULL,
#`Dark Knight`  int  NOT NULL,
#`Grandmas Boy`  int NOT NULL,
#`Shaun Of The Dead`  int NOT NULL,
#`Kill Bill`  int NOT NULL,
#`ZombieLand`  int NOT NULL
#);


#INSERT into Ratings (`Person`,`Gladiator`, `Dark Knight`,`Grandmas Boy`,	`Shaun Of The Dead`,`Kill Bill`,`ZombieLand`)
#VALUES ("Miller",5,5,5,3,5,4),
#("brian",5,5,4,5,5,4),
#("Ariel",4,4,5,4,3,2),
#("Justin",5,4,5,3,5,4),
#("Joia",4,5,1,4,3,5),
#("Stephan",5,5,3,3,3,4),
#("Briana",3,5,5,2,1,2),
#("Rotten_Tomatoes",76,94,16,92,85,90),
#("User_Scores_RT",87,94,85,93,81,90);



#select * from Ratings

```


```{r, echo=FALSE}

```

### Connect to db
+ Run query to select entire table
+edited out the user and pw

```{r}
mydb = dbConnect(MySQL(), user=user, password=my_pass, dbname='movies_db', host='localhost')
user_movie_ratings <- dbGetQuery(mydb, "select * from Ratings ")
user_movie_ratings

```

+ Alternatively, we can use R markdown SQL Chunk to run sql code query after we setup our connection
```{sql connection=mydb}
SELECT * FROM Ratings Where `Shaun Of The Dead` =3

```
## Data Manipulation
+  Add user scores and critics scores from Rotten Tomatoes
+ These scores are on a scale of 1-100
    + Scale them as 1-5
```{r, echo=FALSE} 
library(pander)

#options(knitr.table.format = "latex")
table_2<-read.table(text="
RT_SCORE replacement
0-20 1
21-40 2
41-60 3
61-80 4
81-100 5
" ,  header=TRUE, stringsAsFactors=FALSE)
pander(table_2)

```

```{r}
user_movie_ratings[c(7,9),2:7]<- sweep(user_movie_ratings[c(7,9),2:7],1,c(20,20,20,20,20,20),"/")
user_movie_ratings[c(7,9),2:7] <- ceiling(user_movie_ratings[c(7,9),2:7])
```

## Display how each movie does with all of our "critics"
```{r}

users <- c("Ariel","Brian","Briana","Joia","Justin","Miller","Rotten_Tomatoes","Stephan","User_scores_RT")
movie_list <- list(Gladiator=user_movie_ratings$Gladiator, Dark_Knight=user_movie_ratings$`Dark Knight`,Grandmas_Boy=user_movie_ratings$`Grandmas Boy`,Shaun_Of_The_dead=user_movie_ratings$`Shaun Of The Dead`,Kill_Bill=user_movie_ratings$`Kill Bill`,Zombie_Land =user_movie_ratings$ZombieLand)
                 
dat <- lapply(movie_list, function(x) cbind(x = seq_along(x), y = x))
list_my_names <- names(dat)
lns <- sapply(dat, nrow)
dat <- as.data.frame(do.call("rbind", dat))
dat$group <- rep(list_my_names, lns)
my_vector=user_movie_ratings$Person
my_big_vec <- rep(my_vector,6)
dat$x <- users

library(ggplot2)
ggplot(dat, aes(x = dat$x, y = dat$y, colour = dat$group)) +
    theme_bw() +
    geom_jitter(size=5)+
    theme(axis.text.x=element_text(angle=60, hjust=1))
#matplot(user_movie_ratings$Gladiator)
```
### Observations
+ Movies like Zombie Land and Grandmas Boy have very polarized results
+ Gladiator and Dark Knight are very highly rated
+ Shaun of the dead has the most scores of 3(3)
+ Grouping results by movie will very quickly become difficult to interpret in visual form

## Lets explore the overall tendencies in our critics ratings

```{r}


ggplot(dat, aes(x = dat$x, y = dat$y, colour = dat$x)) +
    theme_bw() +
    geom_jitter(size=7)+
  theme(axis.text.x=element_text(angle=60, hjust=1))
```

## Much Cleaner: lets summarize some results
+ All of the females in the list tend to give more unfavorable ratings
+ Rotten Tomatoes User_scores give every movie on this list a 5 
+ Stephan tends to give the most scores of around 3   
+ Briana gives the lowest scores overall


## Average the scores my friends submitted 
+ Join average friends score with Rotten Tomatoes scores 
+ Added inline comments
```{r}
# Create an empty list and fill with average score of each movie
N <- 7
average_scores <- vector("list", N)
for(i in 2:N) {
    Ps <- mean(user_movie_ratings[,i]) 
    average_scores[[i]] <- Ps
}

# Add average scores to an unlisted vector and round scores to appropriate values
average_scores <- (average_scores[2:7])
friends_avg <- unlist(average_scores)
friends_avg <- ceiling(friends_avg)
names <- "Friends_avg" 
friends_avg <- append(friends_avg, names, after=0)

## Subset old data to get the RT user and critics scores
rotten_tomatoes_subset <- subset(user_movie_ratings,Person== "Rotten_Tomatoes" | Person == "User_Scores_RT")

##combine subset data with friends_avg vector 
comparison_df <- rbind(rotten_tomatoes_subset,friends_avg)
comparison_df

```

## Graph clustering our scores by the the reviewers(RT_users,RT_critics, My_friends) 
```{r}
movie_list <- list(Gladiator=comparison_df$Gladiator, Dark_Knight=comparison_df$`Dark Knight`,Grandmas_Boy=comparison_df$`Grandmas Boy`,Shaun_Of_The_dead=comparison_df$`Shaun Of The Dead`,Kill_Bill=comparison_df$`Kill Bill`,Zombie_Land =comparison_df$ZombieLand)

dat_2 <- lapply(movie_list, function(x) cbind(x = seq_along(x), y = x))
list_my_names <- names(dat_2)
lns <- sapply(dat_2, nrow)
dat_2 <- as.data.frame(do.call("rbind", dat_2))
dat_2$group <- rep(list_my_names, lns)
my_vector_2=comparison_df$Person
my_big_vec <- rep(my_vector_2,6)
dat_2$x <- my_big_vec

ggplot(dat_2, aes(x = dat_2$x, y = dat_2$y, colour = dat_2$x)) +
    theme_bw() +
    geom_jitter(size=10)

```

## That was a fun experiment.  Lets Expore problems with the experimental design
### Discrete rating system
+ Having only 5 ratings to choose from, leaves a very large range for what each rating really means
    + Questions such as "is 5 perfect?" and "what if i want to rate it by 100?" were presented by respondents
+ Presenting these movies together may Bias our results
    + Many of my friends asked questions such as "am I comparing these movies to each other?"  
        + As a large scaled survey, we would want to choose from a larger pool of movies and randomize which selections people were presented.
    + Another Issue brought up was " Am i rating these movies by genre, or overall movie"
        + Perhaps we would want to create strata by genre and present genre based rating surveys

### Our study wasn't blinded
+ Non randomized movie selection
    + Overall poor practice for experimental design.  
    + It is possible our low women scores can be attributed to my gender bias in movie selection
    



