---
title: "Untitled"
author: "Justin Herman"
date: "February 12, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Input raw_data String and attempt to extract numbers with different Regular Expressions

```{r}
library(stringr)
 raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson,Homer5553642Dr. Julius Hibbert"
Simpson_name <- (str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
Simpson_name <- unlist(Simpson_name)
my_numbers <-unlist(str_extract_all(raw.data, "\\d+(\\-| )\\d+"))
my_numbers
get_the_numbers<-str_extract_all(raw.data,"\\d+(\\s|\\-|\\d|\\))+")
get_the_numbers
```

## 1. Create first and Last Simpson_name vectors.  
+ Use str_detect and str_split to populate vectors 

```{r}
Simpson_name
first <- c()
last <- c()
for (x in Simpson_name){
    if  (x=="Simpson,"){
        last <- append(last,"Simpson")
    }   
    else if (str_detect(x,",")==TRUE){
    trials <- unlist(str_split(x,",", n=2))
    first <- append(first,trials[2])
    last <- append(last,trials[1])
    }
    else if (str_detect(x,"\\.")==TRUE){
    trials <- unlist(str_split(x,"\\. ", n=2))
    full_Simpson_name <- trials[2]
    full_Simpson_name <- unlist(str_split(full_Simpson_name,"\\s"))
    first <- append(first,full_Simpson_name[1])
    last <- append(last,full_Simpson_name[2])
    }
    else if (x=="Homer"){
        first <- append(first,x)
    }
    else { trials <- unlist(str_split(x," "))
       # while (is.na(trials)==FALSE){
        first <- append(first,trials[1])
        last <- append(last,trials[2])
     }
}

First_Last_names <- str_c(first,last, sep=" ")

print (First_Last_names)



```

## 2. Construct a Logical vector indicating whether a charachter has a title 
+ Two solutions provided
```{r}
##
rev <- str_detect(Simpson_name,("Rev"))
doctor <- str_detect(Simpson_name,("Dr"))
prefix_check <- rev|doctor
simpson_character_has_prefix_Simpson_name <- cbind(First_Last_names,Simpson_name,prefix_check)
simpson_character_has_prefix_Simpson_name

## Alternative Solution check for Dr. and rev. 
check <- str_detect(Simpson_name,"[DrRev]{2,3}\\.")
check

```
## 3. Constuct a logical Vector indicating wheter a character has a second name
```{r}
two_Simpson_names <- str_replace_all(Simpson_name,"[DRrev]{2,3}\\.","")
#two_Simpson_names
count_Simpson_names <- str_count(two_Simpson_names, "\\w+")
vector_Simpson_namescheck <- str_detect(count_Simpson_names,"3")
```

## Vector returning True if middle name exists
```{r}
vector_Simpson_namescheck
```
## Dataframe summarizing results so far
```{r}
count_Simpson_names <- str_replace_all(count_Simpson_names,"2"," No middle name")
count_Simpson_names <- str_replace_all(count_Simpson_names,"3","middle name")
simpsons_df <- as.data.frame(cbind(First_Last_names,prefix_check,count_Simpson_names))
simpsons_df
    
```
##Question 4
###1  "[0-9]+\\$"))

+ String 1 will find any combination of numbers that ends with a "$" sign
```{r}
string_1 <- "xx44445$xxxx4589$"
unlist(str_extract_all(string_1, "[0-9]+\\$"))


```
### 2 \\b[a-z]{1,4}\\b

+ words of size 1-4 that only use letters 
```{r}
string_2 <- "fourty four fourtyfor fort fa4 ra_p"
unlist(str_extract_all(string_2, "\\b[a-z]{1,4}\\b"))
```
### 3 .*?\\.txt$

+ Reluctant- string that ends with .txt
```{r}
string_3 <-c("Doesn't end with.txt.","does2end with .txt.txt")
unlist(str_extract_all(string_3,".*?\\.txt$"))

```
### 4  \\d{2}/\\d{2}/\\d{4}
+ 2digits / 2 digits/ 4 digits   
+ Looks like a date format although its not restricted to 12 months and 30 days
```{r}
string_4 <- c("12/24/2018", "14/44/4040")
unlist(str_extract_all(string_4,"\\d{2}/\\d{2}/\\d{4}"))

```
### 5 <(.+?)>.+?</\\1>"))

+ tags with any values in between, then a / followed by a back reference to the first tag (.+?) closed with  > 
```{r}
tags <- c("<this_tag_matches_end><any_tag>continue my message 434<tag_3></this_tag_matches_end>")
unlist(str_extract_all(tags, "<(.+?)>.+?</\\1>"))
```

## 9. 

```{r}
secret <- paste("clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo",
                "Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO",
                "d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5",
                "fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr")
uppercased <-unlist(str_extract_all(secret,"[[:upper:]|.]{1}"))

uppercased
message <- str_replace_all(uppercased,"\\.", " ")
yahtzee <- paste(message,collapse="")
yahtzee
```

