---
title: "project"
author: "Justin Herman"
date: "February 17, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Read in text file as string and import Libraries
```{r}
library(stringr)
library(readr)
library(plyr)
library(dplyr)
library(data.table)
mystring <- read_file("tournamentinfo.txt")
```

## Extract Column Names and state names using regular expressions
```{r}

states <- c()
column_names <- unlist(str_extract_all(mystring,"\\w+"))[1:11]
states_2 <- str_split(unlist(str_extract_all(mystring,"[A-Z]+[:blank:]\\|")),"\\s+")
for (x in 1:64){
states <- append(states,states_2[[x]][1])
}
column_names
```
## Extract data in between the | |
+ split data by |
```{r}
trial <- unlist(str_extract_all(mystring,"\\|.*|$"))[1:400]
splited_trial <- str_split(trial[3:130],"\\|")
splited_trial[1:5]
```
## Create blank vectors to populate with chess names, aggregated matches played, player starting rank, and player chess points
+ Use regular expression to populate these vectors
```{r}

chess_names<- c()
matches <- c()
start_ranks <- c()
chess_points <- c()

##If block grabs ranks from splited_trial[[even#'s]] 
##Else catches the rest or splited_trial[[odd#'s]]
for (x in 1:length(splited_trial)){
    if (x%%2==0){
        splited_trial[[x]]<-splited_trial[[x]][2]
        splited_trial[[x]] <- unlist(str_split(unlist(str_extract(splited_trial[[x]][1],":\\s+[[:digit:]]+")),"\\s+"))[2]
        start_ranks <- append(start_ranks,splited_trial[[x]])
    }
    else {
        chess_names <- append(chess_names,splited_trial[[x]][2])
        chess_points <- append(chess_points,splited_trial[[x]][3])
        splited_trial[[x]] <- str_split(splited_trial[[x]] ,"[:blank:]+")
        for (index in 4:10){
            matches <- append(matches,splited_trial[[x]][[index]][2])
        }
    }
}
```

## Trim and Clean data 
```{r}

chess_points <- as.numeric(str_trim(chess_points))
chess_names <- str_trim(chess_names)
matches[matches==""] <- NA
start_ranks <- as.numeric(start_ranks)
matches <- as.numeric(matches) 
chess_names <- str_trim(chess_names)
```

## Split aggreagated matches data into matches by player
+ Map ranking values from another df into my matches by player df 
    + Example matches by player= 39,21,18,14,7,12,4 becomes player_ranks 1605 1469 1564 1574 1501 1519 1372
```{r}
matches_by_player <- split(matches, ceiling(seq_along(matches)/7))
element_1 <- matches_by_player
element_1 <- as_data_frame(element_1)
df_start_ranks <- as_data_frame(start_ranks)
df_start_ranks <- cbind(1:64,df_start_ranks)

temp_df<- as.data.frame(lapply(element_1, function(x)  ifelse(grepl(paste(df_start_ranks$`1:64`, collapse = '|'), x), 
                                                  df_start_ranks$value[match(x, df_start_ranks$`1:64`)], x)))
## find means of the rank of opponents
average_opp_rank <- round(colMeans(temp_df,na.rm=TRUE),0)


```

## Display all the vectors that were created and combine them into a dataframe
```{r}
average_opp_rank <- as.vector(average_opp_rank)
chess_names
start_ranks
chess_points
states
finished_chess_rankings <- as_data_frame(cbind(chess_names,states,chess_points,start_ranks,average_opp_rank))

finished_chess_rankings

write.csv(finished_chess_rankings, file="finished_chess_rankings.csv")
```









