---
title: "Scrape qb salaries"
author: "Justin Herman"
date: "February 28, 2018"
output:
  html_document:
    theme: cerulean
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

## I need Some help here
+ I scrape a website and  return a list of lists with ints that are currently in character form with special symbols
```{r,message=FALSE}
library(stringi)
library(tidyverse)
library(rvest)
library(stringr)
library(rvest)
library(purrr)


qbs <- lapply(paste0('https://overthecap.com/position/quarterback/',2010:2022,"/"),
                function(url){
                    url %>% read_html() %>% 
                        html_nodes('tr:nth-child(2) .mobile_drop+ td , tr:nth-child(3) .mobile_drop+ td , tr:nth-child(4) .mobile_drop+ td , tr:nth-child(5) .mobile_drop+ td , tr:nth-child(1) .mobile_drop+ td') %>% 
                        html_text()
                })

trial_qbs<- qbs 
trial_qbs[[1]]
```

## So here was my workaround for the code
+ Gets me desired result but I had to use a double for loop and it unlists my results.  Unlist is easily fixed here, because most lists are the same size, but in general each list represented a year so I want to avoid it
```{r}
salaries <-c()     

for (x in trial_qbs){
    for (y in x){
        z <- as.integer(str_replace_all(y,"\\$|\\,",""))
        salaries <- append(salaries,z)
        }
    }

str(salaries)




```


## This was my 1st attempt which got me completely lost.  
+ What type are these characters and how can I convert them?
+ Looking at it now I'm guessing i needed to somehow iterate over blank_slate before running replace_all or gsub
    + How can I do that?
```{r}
completely_lost <- str_replace_all(trial_qbs,"\\$|,","")
completely_lost
for (x in seq_along(trial_qbs)){
    trial_qbs[[x]] <- gsub("\\$|,","",trial_qbs[[x]])
}
trial_qbs

trial_qbs<- qbs 


gsub("[^\\d]", "", trial_qbs[[1]], perl=T)

odds <- splited_trial[seq(from=1,to=127,by=2)]
evens <-splited_trial[seq(from=2,to=127,by=2)]
vector_ranks <- lapply(evens, `[[`, 2)
vector_ranks<- unlist(str_split(unlist(str_extract(vector_ranks,":\\s+[[:digit:]]+")),"\\s+"))
vector_ranks[seq(from=2,to=127,by=2)]


```

