---
title: "week7"
author: "Justin Herman"
date: "March 15, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load in Libraries
+ I chose to just add a fake author called ``author B`` to Understanding Power

```{r,message=FALSE}
library(knitr)
library(rvest)
library(RCurl)
library(XML)
library(htmltab)
library(kableExtra)
library(tidyverse)
library(dplyr)
```

## Load in HTML File
+ Being that I loaded the html as a table, this was very easy to load
+ Should be reproducible as it loads from my github
```{r}
url <- "https://raw.githubusercontent.com/justinherman42/Justin-Data-607/master/tidy_data_week4/Politicalbooks.html"
my_books<- htmltab(doc = url, which = "/html/body/table")
kable(my_books)

```

## Load in XML File
+ Using Xmlparse
    + I initially received alot of errors attempting to load from github, found workarounds on stack overflow
    + Tidy data 
```{r}
fileURL <- "https://raw.githubusercontent.com/justinherman42/Justin-Data-607/master/tidy_data_week4/Politicalbooks2.xml"
books_xml <- getURL(fileURL,ssl.verifypeer = FALSE)
books_xml %>% 
    xmlParse(.,useInternal = TRUE) %>% 
    xmlToList(.) %>%  
    plyr::ldply(., data.frame) %>% 
    select(-.id) %>%
    mutate(Co_author=Author.1) %>%
    select(-Author.1) %>% 
    kable(.)

#books_xml <- xmlParse(books_xml ,useInternal = TRUE)
#xL <- xmlToList(books_xml)
#kable(ldply(xL, data.frame))
```

## Alternate example of loading in xml
+ Couldn't get this example to work with git hub
+ There are no 2nd authors in this example
+ the df needed to me transposed 
```{r}
xml.url <- "file:///C:/Users/justin/Documents/GitHub/Justin-Data-607/tidy_data_week4/Politicalbooks.xml"
xmlfile <- xmlTreeParse(xml.url)
class(xmlfile)
xmltop = xmlRoot(xmlfile)
print(xmltop)[1:2]
plantcat <- xmlSApply(xmltop, function(x) xmlSApply(x, xmlValue))
plantcat_df <- t(as_data_frame(plantcat))
colnames(plantcat_df) <-c("Book", "Author", "Publication Date","Publisher")
kable(plantcat_df)


```

## Load in JSON file
```{r}
library(curl)
library(rjson)
library(jsonlite)
json_file <- 'https://raw.githubusercontent.com/justinherman42/Justin-Data-607/master/tidy_data_week4/Politicalbooks.json'
text <- readLines(curl(json_file))
text %>% 
    jsonlite::fromJSON(.,flatten = TRUE ) %>% 
    kable(.)
    


#text
#dd  <-  as.data.frame(t(matrix(unlist(json_data), nrow=4)))
#colnames(dd) <- c("Book", "Author", "Publication Date","Publisher")
#kable(dd)
```

## 
```{r}
# require(RCurl)
# require(XML)
# webpage <- getURL("https://raw.githubusercontent.com/justinherman42/Justin-Data-607/master/tidy_data_week4/Politicalbooks.xml")
# webpage <- readLines(tc <- textConnection(webpage)); close(tc)
# pagetree <- htmlTreeParse(webpage, error=function(...){}, useInternalNodes = TRUE)
# x <- xpathSApply(pagetree, "/html/body", xmlValue)  
# x <- strsplit(x, ",")
# dd_2  <-  as.data.frame(t(matrix(unlist(x), nrow=4)))
# colnames(dd_2) <- c("Book", "Author", "Publication Date","Publisher")
# kable(dd_2)
# x
# 
# 
# 
# 
# require(RCurl)
# require(XML)
# webpage <- getURL("https://raw.githubusercontent.com/justinherman42/Justin-Data-607/master/tidy_data_week4/Politicalbooks2.xml")
# webpage <- readLines(tc <- textConnection(webpage)); close(tc)
# pagetree <- htmlTreeParse(webpage, error=function(...){}, useInternalNodes = TRUE)
# x <- xpathSApply(pagetree, "/html", xmlValue)  
# x <- strsplit(x, ",")
# dd_2  <-  as.data.frame(t(matrix(unlist(x), nrow=4)))
# colnames(dd_2) <- c("Book", "Author", "Publication Date","Publisher")
# kable(dd_2)
# 
# 
# rm (list=ls())




```