---
title: "project_2_c"
author: "Justin Herman"
date: "March 11, 2018"
output:
  html_document:
    theme: cerulean
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(kableExtra)
library(knitr)
library(stringr)
library(readr)
library(plyr)
library(dplyr)
library(data.table)
library(readxl)
require(xlsx)
library(rJava)
library(xlsx)

```


# Third Dataset- Gapminder
+ Burcu recommended the Gapminder data website
+ [link to source](https://www.gapminder.org/data/)


## War or food?
+ Are nations dealing with war more food insecure?
+ How do state expenditures differ on war versus food?

## Load in third dataset 
+ File 1 is a file with death rates out of 100k from war  [link to source](https://docs.google.com/spreadsheets/d/15kseoFj-w39TvMl_Ije-C2TZy-DKYyECsF-uGdec0z8/pub)
+ File 2 is a measurement of daily calories consumption    [link to source](https://docs.google.com/spreadsheets/d/14G6CjF6NblTGf6kkQclpXp3XZ3D4Nkw1I92DB4fOjXo/pub)

+ File 3 is GDP War expenditure %                           [link to source](https://docs.google.com/spreadsheets/d/1HXH3ncdQLFpOXVNdC2nkPCSGz_perkfPcBeU1e2HF24/pub)

+ File 4 is GDP Agricultural Expenditure %                   [link to source](https://docs.google.com/spreadsheets/d/16JAXDjaKcyaEfd6sO_pRC892NG8VAk6moyw3Gl50-Ys/pub)

```{r}
delete.na <- function(DF, n=0) {
  DF[rowSums(is.na(DF)) <= n,]
}


War<- read.xlsx("War.xlsx", header=TRUE,1,)
food_security <- read.xlsx("indicator food_consumption.xlsx", header=TRUE,1,)
Agricultural_expend<- read.xlsx("Agricultural_expenditure.xlsx", header=TRUE,1,)
military_expend <- read.xlsx("military_expenditure.xlsx", header=TRUE,1,)

dim(food_security)
dim(military_expend)
dim(Agricultural_expend)
dim(War)
```

## DF Preparation 
+ Eliminate rows with substantial NA's
+ Convert columns to numeric 
+ Rename columns to avoid duplication of Column names for eventual Join
+ Filter df's for common shared year range of 1988-2007
```{r}
military_expend <- delete.na(military_expend, 20)
Agricultural_expend <- delete.na(Agricultural_expend, 15)
food_security <- delete.na(food_security, 20)
War <- as_data_frame(lapply(War, function(x) if(is.numeric(x)) round(x,2) else x))
military_expend <- as_data_frame(lapply(military_expend, function(x) if(is.numeric(x)) round(x,2) else x))
War <- War[2:4]
colnames(War) <- c("Country", 'war_2002','war_2004')
colnames(military_expend) <-  c("Country",paste(1988:2011,'Military_expend'))
colnames(food_security) <- c("Country",paste (1961:2007,"food_security"))
colnames(Agricultural_expend) <- c("Country",paste(1961:2011,"Agricultural_expend"))
Agricultural_expend <- Agricultural_expend[c(1,29:48)]
food_security <- food_security[c(1,29:48)]
military_expend <- military_expend[c(1:21)]


merged_df_3<- inner_join(War,military_expend) %>%
    inner_join(.,food_security) %>% 
    inner_join(.,Agricultural_expend)  
dim(merged_df_3)

```


## Exploratory Analysis of Death from wars data 
+ Explore war death df
```{r}
merged_df_3$war_2002-merged_df_3$war_2004


merged_df_3[1:3] %>% 
    mutate(.,average_deaths=(war_2002+war_2004)/2) %>% 
    arrange(.,average_deaths) %>%  
    kable(.)
 
```

## Tidy dataset
+ Add average deaths from war column
+ Create a categorical factor level of ``deaths from war``
    + >0 No_Deaths 
    + >0-1 Low_Deaths 
    + >1-10 Moderate_Deaths 
    + >10+ High_Deaths

```{r}
food_war <- merged_df_3 %>% 
    gather(Wars,Values,4:63) %>% 
    separate(.,Wars,into= c("Year", "Type"),sep=" ") %>% 
    spread(Type,Values) %>% 
    mutate(.,average_deaths=(war_2002+war_2004)/2) %>% 
    select(-c(war_2002,war_2004)) %>% 
    mutate(Deaths_from_war=cut(average_deaths, breaks=c(-1,0,1,10,Inf), labels= c("No_Deaths", "Low_Deaths", "Moderate_Deaths", "High_Deaths")))
   
    

kable(food_war[1:50,])
```

## Summary Explorations
```{r}

food_war <-food_war %>% 
    dplyr::group_by(Deaths_from_war,Year)
agricultural_spending_by_year <- food_war %>% 
    dplyr::summarize(agri_exp_mean=mean(Agricultural_expend, na.rm = TRUE)) 
food_war %>% 
    dplyr::summarize(Military_Spending_Pct_GDP_By_Year=mean(Military_expend, na.rm = TRUE)) %>% 
    kable(.)
food_security_by_year <- food_war %>% 
    dplyr::summarize(my_mean=mean(food_security, na.rm = TRUE))

food_war <-food_war %>% 
    dplyr::group_by(Deaths_from_war)
food_war %>% 
    dplyr::summarize(Mean_Agricultural_Spending_Pct_GDP=mean(Agricultural_expend, na.rm = TRUE)) %>% 
    kable(.)
food_war %>% 
    dplyr::summarize(Mean_Military_Spending_Pct=mean(Military_expend, na.rm = TRUE)) %>% 
    kable(.)
food_war %>% 
    dplyr::summarize(mean_Caloric_consuimption=mean(food_security, na.rm = TRUE)) %>% 
    kable(.)
  
```

## Warnings about observations
+ My categories aren't of even sizes so my clustering could effect my observations and lead to large impact from outliers
+ I don't have an overall GDP per country, this could make a big difference in relative % GDP expenditures
+ Tentative Observations
    + Areas with the most death from wars seem to spend more of their GDP on agricultural expenditures
    + Areas with the most deaths don't spend much on military expenditures, and low death areas actually spend more than high death areas 
    + High and moderate death areas are significantly more food insecure. Calorie consumption is about 30% lower in these areas




## Data Visualization 
```{r}
food_security_by_year

food_security_by_year %>% 
    ggplot(.,aes(x=Year,y=my_mean))+
    geom_bar(aes(fill = Deaths_from_war), position = "dodge", stat = "identity")+
    theme(axis.text.x=element_text(angle=45,hjust=1))
```

+ Caloric consumption seems rather Flat in all the categories over the last three decades


```{r}
agricultural_spending_by_year %>% 
    ggplot(.,aes(x=Year,y=agri_exp_mean))+
    geom_bar(aes(fill = Deaths_from_war), position = "dodge", stat = "identity")+
    theme(axis.text.x=element_text(angle=45,hjust=1))
```

+ As a percent of GDP agricultural expenditure has decreased across all categories over last three decades
+ Area's with higher casualties of war seem to have made large cuts to expenditures


## Most interesting graphs and observations
+ A handful of outliers were removed for better visualization 

```{r}



food_war %>% 
    ggplot(.,aes(x=Agricultural_expend,y=food_security,shape=Deaths_from_war,color=Deaths_from_war))+
    geom_point()
```

### Graph 1 Agricultural Exp vs Food Security
+ Two Clear Clusters develop in the first graph
    + Areas with high death rates seem to be the least food secure and simultaneously spend the most on agricultural subsidy
    + Areas with low death rates spends the least on food subsidy, and are the most food secure


```{r}
food_war %>% 
    filter(.,Military_expend <20) %>% 
    ggplot(.,aes(x=Military_expend,y=food_security,shape=Deaths_from_war,color=Deaths_from_war))+
    geom_point()
        




```


### Graph 2 Military Exp vs Food Security 
+ Areas with low death rates seem to spend the most on the Military


