---
title: "project 2a"
author: "Justin Herman"
date: "March 8, 2018"
output:
  html_document:
    theme: cerulean
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load in libraries


```{r, message=FALSE}
library(tidyverse)
library(kableExtra)
library(knitr)
library(stringr)
library(readr)
library(plyr)
library(dplyr)
library(data.table)
library(readxl)
require(xlsx)
library(rJava)
library(xlsx)
rm (list=ls())
```

## Tidy dataset 1: Movies database
+ Posted by Betsy
+ [link to source](https://docs.google.com/spreadsheets/d/1K_a87dxLQR91GwVcqkqZK9kjxcYDl1tbK9EX9WvC7bo/edit#gid=601448852)

## Load in Data from xlsx file
+ Dataframe requires tidying
    + Join individual movies into one movies column
    + Separate timestamp by day and time
        + Could perform some interesting time series analysis on the effects of time of day of review
    + Omit NA Values( the file read in 100+ extra rows of NA) 
    + Convert 0's to NA assuming that 0's meant the person didn't see the movie
        + This could be a bad assumption and perverse the data
        + Two reviewers didnt review a single movie, so I will filter them out
```{r}
movies <- as.data.frame(read.xlsx("Movie Ratings.xlsx", header=TRUE,1))
#tail(movies)
#dim(movies)

movies%<>% 
    gather(Movies,Movie_rating,3:12) %>% 
    select(Timestamp,Reviewer=Name,Movies,Movie_rating,Number.of.Movies.Seen) %>%   
    separate('Timestamp',c("day","time")," ") %>% 
    na.omit() %>% 
    arrange(Reviewer) %>% 
  #  mutate(Reviewer==Name) %>% 
    replace(.==0,NA) %>% 
    filter(Movie_rating>0)


    
movies %>% 
    dplyr::group_by(Movies) %>% 
    dplyr::summarise(Movie_rating=round(mean(Movie_rating,na.rm=TRUE),2)) %>% 
    kable(.)
```

## data Visualization
+ Display average rating by user
+ Display average rating by movie
+ Subset and display active critics(people who have seen over 5 movies) average ratings
+Display average rating by movie from active critics
```{r}

movies %>%  
    dplyr::group_by(Reviewer) %>% 
    dplyr::summarise(Average_Review=round(mean(Movie_rating,na.rm=TRUE),2)) %>% 
    arrange(desc(Average_Review)) %>% 
    ggplot(., aes(x=Reviewer , Average_Review)) + 
    geom_bar(aes(fill = Reviewer), position = "dodge", stat = "identity")+
    coord_flip()+
    labs(title="Average Rating By User")

movies %>%  
    dplyr::group_by(Movies) %>% 
    dplyr::summarise(Average_Review=round(mean(Movie_rating,na.rm=TRUE),2)) %>% 
    arrange(desc(Average_Review)) %>% 
    ggplot(., aes(x=Movies , Average_Review)) + 
    geom_bar(aes(fill = Movies), position = "dodge", stat = "identity")+
    theme(axis.text.x=element_text(angle=45,hjust=1))+
    labs(title="Average Rating By Movie")
    
 movies %>%  
    dplyr::group_by(Reviewer) %>% 
    filter(Number.of.Movies.Seen>5) %>% 
    dplyr::summarise(Average_Review=round(mean(Movie_rating,na.rm=TRUE),2)) %>% 
    ggplot(., aes(x=Reviewer , Average_Review)) + 
    geom_bar(aes(fill = Reviewer), position = "dodge", stat = "identity")+
    theme(axis.text.x=element_text(angle=45,hjust=1))+
    labs(title="Active Critics")
   
movies %>%  
    dplyr::group_by(Movies) %>% 
    dplyr::summarise(Average_Review=round(mean(Movie_rating,na.rm=TRUE),2)) %>% 
    arrange(desc(Average_Review)) %>% 
    ggplot(., aes(x=Movies , Average_Review)) + 
    geom_bar(aes(fill = Movies), position = "dodge", stat = "identity")+
    coord_flip()+
    labs(title="Active Critics Average Rating By Movie")




```

    
