---
title: "Project_2_b"
author: "Justin Herman"
date: "March 11, 2018"
output:
  html_document:
    theme: cerulean
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message=FALSE}
library(tidyverse)
library(kableExtra)
library(knitr)
library(stringr)
library(readr)
library(plyr)
library(dplyr)
library(data.table)
library(readxl)
require(xlsx)
library(rJava)
library(xlsx)
```


## World Bank Info For Infant Mortality And Tuberculosis
+ **Binish recommended this dataset website**
+ [link to source](http://apps.who.int/gho/data/node.country.country)
    + **Update I believe this link went dead.  I'll send the csv's with upload**
+ **Randomly chose 4 countires with the letter `U`**
+ **Read in files individually/with functional**
```{r}

#Manually Load in CSV
Uganda<- read.csv("uganda.csv",na.strings=c("","NA"))
Ukraine <- read.csv("Ukraine.csv",na.strings=c("","NA"))
United_kingdom <- read.csv("United_Kingdom.csv",na.strings=c("","NA"))
United_states <- read.csv("United_states.csv",na.strings=c("","NA"))


# Functional to load csv
files <- c("Uganda.csv", "Ukraine.csv", "United_Kingdom.csv", "United_states.csv")
temp = files
myfiles <-lapply(temp,read.csv)
```


## Join On Shared Row Index Values(Survey Questions)
+ **This will decrease our sample to only those questions that are answered in every dataset**
```{r,message=FALSE,warning=FALSE}

merged_df<- inner_join(United_states,United_kingdom) %>% 
    inner_join(.,Ukraine) %>% 
    inner_join(.,Uganda)

# With Functional data
merged_df_2<- inner_join(myfiles[[1]],myfiles[[2]]) %>% 
    inner_join(myfiles[[3]]) %>%   
    inner_join(myfiles[[4]])

```

```{r}
kable(merged_df_2[1:15,])
```



## Tidy Dataset

+  **Gather all country data into one column**
+  **Create a year column to map to the df**
+  **Spread the questions into columns with their values**

```{r}

getting_tidy <- merged_df[-(1),] %>% 
    gather(United_states,Values,2:57) %>% 
    mutate(Year=rep(unlist(lapply(2015:2002,function(x) rep(x,23))),4)) %>% 
    spread(X, Values) 
kable(getting_tidy[1:10,])

```    

## Get Rid Of Special Characters ``]`` ``[``  ``-``    
+ **Create standardized spacing in each value of DF  for next pipe of tidying**
+ **Use pipe to Tidy ``Mean[min-max]`` columns into Mean,Min and Max columns for respected columns**
```{r}

getting_tidy <-  as_data_frame(lapply(getting_tidy,function(x){ str_replace_all(x,"\\]|\\["," ")}))
getting_tidy <-  as_data_frame(lapply(getting_tidy,function(x){ str_replace_all(x,"-","  ")}))



final_df <- getting_tidy %>% 
    separate(.,`Incidence of tuberculosis (per 100 000 population per year)`,into=c('Mean_Incidence_Tuberculosis_100,000','Min_Tuberculosis',"Max_Tuberculosis")) %>% 
    separate(.,`Antiretroviral therapy coverage among people with HIV infection eligible for ART according to 2010 guidelines (%)`,into=c('Mean_Antiretroviral_coverage','Min_Antiretroviral_coverage','Max_Antiretroviral_coverage')) %>% 
    separate(.,`Infant mortality rate (probability of dying between birth and age 1 per 1000 live births)`,into=c('Mean_Infant mortality_Rate/1000','Min_Infant_Mortality rate',"Max_Infant_Mortality_Rate"),sep= "  ") %>% 
    separate(.,`Under-five mortality rate (probability of dying by age 5 per 1000 live births)`,into=c('Mean_Under-five_Mortality_Rate/1000','Min_Under-five_Mortality_Rate',"Max_Under-five_Mortality_Rate"), sep="  ") %>% 
    separate(.,`Tuberculosis treatment coverage`,into=c('Mean_Tuberculosis_Coverage','Min_Tuberculosis_Coverage',"Max_Tuberculosis_Coverage"))


```


##  More Piping
+ **First column needs to be renamed to ``Country`` column **
+ **Eliminate the ``.integers`` after the values in ``Country`` column**
+ **Get rid of some columns**
+ **Trim DF**
+ **Convert columns to numeric for analysis**

```{r}
final_df %<>% 
    plyr::rename(.,c('United_states'= 'Country')) %>% 
    mutate(Country=rep(c("Uganda","Ukraine","United_Kingdom","United_States"),each=14)) %>% 
    select(colnames(.)[c(1,2,10,11,12,13,14,15,16,17,21,22,23,30,31,32,33,34,35)]) %>% 
    arrange(.,`Country`)  

final_df[] <- lapply(final_df,function(x){ str_trim(x)}) 
final_df[3:19] <- lapply(final_df[3:19],function(x) as.numeric(as.character(x))) 
final_df <- as_data_frame(final_df)
kable(final_df)
```  

## Visual Exploratory Analysis Of Infant Mortality
+ **GNP By Year**
+ **Infant Mortality Rate by year**
+ **Numerical vector of Infant Mortality Rate by year**

```{r}
final_df %>%
    mutate(GNP=as.numeric(`Gross national income per capita (PPP int. $)`)) %>% 
    dplyr::group_by(Country) %>% 
    ggplot(., aes(x=Year ,y=GNP))+ 
    geom_bar(aes(fill = Country), position = "dodge", stat = "identity")+
    theme(axis.text.x=element_text(angle=45,hjust=1))+
    labs(title="GNP By Year")
    
final_df %>%
    dplyr::group_by(Country) %>% 
    ggplot(., aes(x=Year ,y=`Mean_Infant mortality_Rate/1000`))+ 
    geom_bar(aes(fill = Country), position = "dodge", stat = "identity")+
    theme(axis.text.x=element_text(angle=45,hjust=1))

final_df_display <- final_df %>% 
    arrange(Country,Year) %>% 
    select(.,Country,`Mean_Infant mortality_Rate/1000`)

```

+ **Uganda seems to have reduced it's infant mortality rate by over 50%**

## Numerical Display Of Infant Mortality
```{r}
## Numerical display     
kable(final_df_display)
```

## Display Of Infant Mortality With The 3 Samples Closer In Scale

```{r}
final_df %>%
    dplyr::filter(., Country %in% c("United_Kingdom", "Ukraine","United_States")) %>% 
    dplyr::group_by(Country) %>% 
    ggplot(., aes(x=Year ,y=`Mean_Infant mortality_Rate/1000`))+ 
    geom_bar(aes(fill = Country), position = "dodge", stat = "identity")+
    theme(axis.text.x=element_text(angle=45,hjust=1))


final_df
```

+ **Ukraine has seen a reduction of around 40%**
+ **US and UK have also seen a reduction, however UK seems to have a more substantial reduction(30%) over the US reduction of (17%)**


## Look At Tuberculosis

```{r}
final_df


final_df %>%
    dplyr::filter(., Country %in% c("Uganda", "Ukraine","United_Kingdom","United_States")) %>% 
    dplyr::group_by(Country) %>% 
    ggplot(., aes(x=Year ,y=`Mean_Incidence_Tuberculosis_100,000`))+ 
    geom_bar(aes(fill = Country), position = "dodge", stat = "identity")+
    theme(axis.text.x=element_text(angle=45,hjust=1))
```

+ **Uganda seems to have seen some moderate improvement over the past 13 years in tuberculosis, but the rate is relatively high still**
+ **Ukraine seems to have had a similar improvement**

## Look At Tuberculosis In Heavily Industrialized Nations( UK, US)

```{r}
final_df %>%
    dplyr::filter(., Country %in% c("United_Kingdom","United_States")) %>% 
    dplyr::group_by(Country) %>% 
    ggplot(., aes(x=Year ,y=`Mean_Incidence_Tuberculosis_100,000`))+ 
    geom_bar(aes(fill = Country), position = "dodge", stat = "identity")+
    theme(axis.text.x=element_text(angle=45,hjust=1))
```

+ **England appears to have made some moderate improvement in the past several years**
+ ***Tuberculosis in the United states is almost non-existant and it appears it is decreasing**




